---
title: "Project2"
output:
  html_document: default
  pdf_document: default
---

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(jpeg)
library(png)
library(dplyr)
library(tibble)
library(ggplot2)

```




# __Question1__:Do republican and democratic state have different number of infected people by covid? 
# find the linear equation and compare them 



This is the plot showing  total percentage of people who infected by Covid19 until 2020.11.29 and I can see more than 30% of people in Delaware are infected by covid
but is it shown that most of republican states have higher rate of infection compared to democratic states.



```{r echo=FALSE}
covid=read.csv("covid.csv")
rcovid=covid[,-c(1,2,4)] %>% group_by(State) %>% summarise_all(list(mean))
rcovid=rcovid[-1,]
for( i in 1:51){
rcovid[i,3:679]=rcovid[i,3:679]/as.numeric(rcovid[i,2])
}
rcovid2=rcovid[,-2]
date=seq(as.Date('2020/01/23', '%Y/%m/%d'), +   as.Date('2021/11/29', '%Y/%m/%d'),+1)
last=data.frame(x=rcovid$State,y=rcovid2$X2021.11.29)
last$z=c("republic","republic","republic","republic","democratic","democratic","democratic","democratic","democratic","neutral","republic","democratic","neutral","republic","democratic","republic","republic","republic","republic","democratic","democratic","democratic","democratic","democratic","republic","republic","republic","republic","republic","republic","democratic","democratic","democratic","democratic","democratic","neutral","republic","democratic","democratic","democratic","republic","republic","republic","republic","republic","democratic","democratic","democratic","democratic","republic","republic")
ggplot() +  geom_point(mapping=aes(x=x, y=y, color=z), data=last)+xlab("State")+ ylab("percantage") + ggtitle("cumulative percantage until 2021.11.29") 


```





I found in  Google that shows summary of results of the 2008, 2012, 2016, and 2020 presidential elections by state

red shows states carried by Republicans in all four elections   

orange shows states carried by Republicans in three of the four elections  

purple shows states carried by  each party twice in the four elections  

sky blue shows states carried by the Democrats in three of the four elections 

blue shows the states carried by  Democrats in all four elections

```{r echo=FALSE}
my_image = readPNG("state.png")
plot(1:2, type='n', main="", xlab="x", ylab="y")
lim <- par()
rasterImage(my_image, 
            xleft=1, xright=2, 
            ybottom=1, ytop=2)
grid()
```


and shown by the plot,

democratic states are:WA,OR,CA,NV,HI,CO,NM,MN,WI,IL,MI,PA,NY,NJ,CT,MA,RI,ME,VT,NH,VA,MD,DE

republican states are:ID,MT,WY,UT,AZ,AK,ND,SD,NE,KS,OK,TX,LA,AR,MO,KY,TN,MS,AL,GA,SC,NC,WV,IN

and IA,FL,OH are neutral states



I found a .csv file in goggle that shows Cumulative infection rate per every cities in U.S 
```{r echo=FALSE}
covid=read.csv("covid.csv")
str(covid)
```
and this is structure of original data. 

At first, I added all the people infected by Covid, and divided by population of each cities to find out percentage. 
and I classified  all the cities by states and dates by once in a month.


```{r echo=FALSE}

coviddata=covid[,c("State","StateFIPS","population"
,"X2020.01.23","X2020.02.23","X2020.03.23","X2020.04.23"
,"X2020.05.23","X2020.06.23","X2020.07.23","X2020.08.23","X2020.09.23","X2020.10.23"
,"X2020.11.23","X2020.12.23","X2021.01.23","X2021.02.23","X2021.03.23","X2021.04.23","X2021.05.23"
,"X2021.06.23","X2021.07.23","X2021.08.23","X2021.09.23","X2021.10.23","X2021.11.23")]
percentage=100
coviddata2 = coviddata %>% group_by(State) %>% summarize(X2020.01.23=sum(X2020.01.23,na.rm=T),                                                                                   X2020.02.23=sum(X2020.02.23,na.rm=T),
                                                         X2020.03.23=sum(X2020.03.23,na.rm=T),
                                                         X2020.04.23=sum(X2020.04.23,na.rm=T),
                                                         X2020.05.23=sum(X2020.05.23,na.rm=T),
                                                         X2020.06.23=sum(X2020.06.23,na.rm=T),
                                                         X2020.07.23=sum(X2020.07.23,na.rm=T),
                                                         X2020.08.23=sum(X2020.08.23,na.rm=T),
                                                         X2020.09.23=sum(X2020.09.23,na.rm=T),
                                                         X2020.10.23=sum(X2020.10.23,na.rm=T),
                                                         X2020.11.23=sum(X2020.11.23,na.rm=T),
                                                         X2020.12.23=sum(X2020.12.23,na.rm=T),
                                                         X2021.01.23=sum(X2021.01.23,na.rm=T),
                                                         X2021.02.23=sum(X2021.02.23,na.rm=T),
                                                         X2021.03.23=sum(X2021.03.23,na.rm=T),
                                                         X2021.04.23=sum(X2021.04.23,na.rm=T),
                                                         X2021.05.23=sum(X2021.05.23,na.rm=T),
                                                         X2021.06.23=sum(X2021.06.23,na.rm=T),
                                                         X2021.07.23=sum(X2021.07.23,na.rm=T),
                                                         X2021.08.23=sum(X2021.08.23,na.rm=T),
                                                         X2021.09.23=sum(X2021.09.23,na.rm=T),
                                                         X2021.10.23=sum(X2021.10.23,na.rm=T),
                                                         X2021.11.23=sum(X2021.11.23,na.rm=T),
                                                         population=sum(population,na.rm=T)
                                                        )  %>% mutate(a2020.01.23=X2020.01.23/population*percentage,
                                                                      a2020.02.23=X2020.02.23/population*percentage,
                                                                      a2020.03.23=X2020.03.23/population*percentage,
                                                                      a2020.04.23=X2020.04.23/population*percentage,
                                                                      a2020.05.23=X2020.05.23/population*percentage,
                                                                      a2020.06.23=X2020.06.23/population*percentage,
                                                                      a2020.07.23=X2020.07.23/population*percentage,
                                                                      a2020.08.23=X2020.08.23/population*percentage,
                                                                      a2020.09.23=X2020.09.23/population*percentage,
                                                                      a2020.10.23=X2020.10.23/population*percentage,
                                                                      a2020.11.23=X2020.11.23/population*percentage,
                                                                      a2020.12.23=X2020.12.23/population*percentage,
                                                                      a2021.01.23=X2021.01.23/population*percentage,
                                                                      a2021.02.23=X2021.02.23/population*percentage,
                                                                      a2021.03.23=X2021.03.23/population*percentage,
                                                                      a2021.04.23=X2021.04.23/population*percentage,
                                                                      a2021.05.23=X2021.05.23/population*percentage,
                                                                      a2021.06.23=X2021.06.23/population*percentage,
                                                                      a2021.07.23=X2021.07.23/population*percentage,
                                                                      a2021.08.23=X2021.08.23/population*percentage,
                                                                      a2021.09.23=X2021.09.23/population*percentage,
                                                                      a2021.10.23=X2021.10.23/population*percentage,
                                                                      a2021.11.23=X2021.11.23/population*percentage)
coviddata2=coviddata2 %>% select(State,a2020.01.23,a2020.02.23,a2020.03.23,a2020.04.23,a2020.05.23,a2020.06.23,a2020.07.23,
                      a2020.08.23,a2020.09.23,a2020.10.23,a2020.11.23,a2020.12.23,a2021.01.23,a2021.02.23,a2021.03.23,
                      a2021.04.23,a2021.05.23,a2021.06.23,a2021.07.23,a2021.08.23,a2021.09.23,a2021.10.23,a2021.11.23)

coviddata2=coviddata2[-1,]

# found the percentage of population who infected by covid


republic=coviddata2 %>% filter(State %in% c("AZ","AL","AK","AR","ID","IN","KS","KY","MS","MT","NE","ND","OK","SC","SD","UT","VA","WV","WY","TX","LA","MO","TN","GA","NC"))
Democratic=coviddata2 %>% filter(State %in% c("CA","CO","CT","DC","DE","GA","HI","IL","MA","MD","ME","MI","MN","NH","NJ","NM","NV","NY","OR","PA","RI","VT","VA","WA","WI"))

#FL,OH,IA are considered swing state so I'll not consider

meanrepublic=republic %>% summarise(m20.01.23=mean(a2020.01.23),m20.02.03=mean(a2020.02.23),m20.03.23=mean(a2020.03.23),m20.04.23=mean(a2020.04.23),
                       m20.05.23=mean(a2020.05.23),m20.06.23=mean(a2020.06.23),m20.07.23=mean(a2020.07.23),m20.08.23=mean(a2020.08.23),
                       m20.09.23=mean(a2020.09.23),m20.10.23=mean(a2020.10.23),m20.11.23=mean(a2020.11.23),m20.12.23=mean(a2020.12.23),
                       m21.01.23=mean(a2021.01.23),m21.02.23=mean(a2021.02.23),m21.03.23=mean(a2021.03.23),m21.04.23=mean(a2021.04.23),
                       m21.05.23=mean(a2021.05.23),m21.06.23=mean(a2021.06.23),m21.07.23=mean(a2021.07.23),m21.08.23=mean(a2021.08.23),
                       m21.09.23=mean(a2021.09.23),m21.10.23=mean(a2021.10.23),m21.11.23=mean(a2021.11.23))

meandemocra= Democratic %>% summarise(m20.01.23=mean(a2020.01.23),m20.02.03=mean(a2020.02.23),m20.03.23=mean(a2020.03.23),m20.04.23=mean(a2020.04.23),
                                    m20.05.23=mean(a2020.05.23),m20.06.23=mean(a2020.06.23),m20.07.23=mean(a2020.07.23),m20.08.23=mean(a2020.08.23),
                                    m20.09.23=mean(a2020.09.23),m20.10.23=mean(a2020.10.23),m20.11.23=mean(a2020.11.23),m20.12.23=mean(a2020.12.23),
                                    m21.01.23=mean(a2021.01.23),m21.02.23=mean(a2021.02.23),m21.03.23=mean(a2021.03.23),m21.04.23=mean(a2021.04.23),
                                    m21.05.23=mean(a2021.05.23),m21.06.23=mean(a2021.06.23),m21.07.23=mean(a2021.07.23),m21.08.23=mean(a2021.08.23),
                                    m21.09.23=mean(a2021.09.23),m21.10.23=mean(a2021.10.23),m21.11.23=mean(a2021.11.23))




numberofmonth=1:23
repub=as.numeric(meanrepublic[1,])
demo=as.numeric(meandemocra[1,])

re=data.frame(numberofmonth,repub)
de=data.frame(numberofmonth,demo)
```

Now from the data I filtered by _republican_ state _democratic_ state, I found a linear equation and 
tried to plot them together to see the difference of infection.

```{r echo=FALSE}
republicequation=lm(numberofmonth~repub,re);republicequation
democraticequation=lm(numberofmonth~demo,de);democraticequation
plot(de,ylim=c(0,17),type="p",xlab="month started from 2020.01.23",ylab="percentage of total population infected by covid",main="United States",col="blue",cex=.5)
abline(lm(de$demo~de$numberofmonth),col="blue")
par(new=TRUE)
plot(re,ylim=c(0,17),type="p",xlab="month started from 2020.01.23",ylab="percentage of total population infected by covid",main="United states",col="red",cex=.5)
abline(lm(re$repub~re$numberofmonth),col="red")
```




## __first conclusion__: It is clear by the plot that over 15% percent of people in republican states were infected by covid untill 2021.11.23 which is over about 3% than people in democratic states.






# __Question2__:Why does number of people infected by covid started decreasing at certain moment? 
# I guess it's because of vaccination, lets see my prediction is right and compare before vaccination and after vaccination.



After finishing my first question, I was curious why number of people infected by covid looked decreased after 12~13 month.
I thought it was because of the vaccine so approximate time people are benefited from vaccinated is about  "2021.01.23"
by the plot I first made. 

By the Goggle, since first vaccination started "2020.12.14" it looks quite reasonable to agree with my thought because vaccine works after 3 weeks. 


Because I only need data of people who infected by Covid19 in __each days__ before vaccination and after vaccination, I used __non-cumulative data__ regardless of State. It only shows every day information that contains __average number of people infected in cities__.  


```{r echo=FALSE}

nocumlative=covid

for (i in 682:8){
  nocumlative[,i]=nocumlative[,i]-nocumlative[,i-1]
}

nocumlative=nocumlative[,8:682]

sum=colMeans(nocumlative,na.rm=T)

date=seq(as.Date('2020/01/25', '%Y/%m/%d'), +   as.Date('2021/11/29', '%Y/%m/%d'),+1)

sum2=data.frame(x=date,y=as.numeric(sum))

sum2$z=c(rep("before vaccine",365),rep("after vaccine",310))

ggplot(sum2) + geom_line(aes(x = x, y = y, color=z), cex = 0.8, show.legend = T) + ggtitle("Covid infection") +xlab("Date") + ylab("average number of people infected in each cities")

```


## __second conclusion__: as shown by the plot, after "2021.01.23" number of people infected started to decrease because of vaccination.




